PANCHANGE_descriptives

### 2020-09-15 PANCHANGE_analysis.Rmd
split of scripts

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "/four_cohorts.rds"))
dim(dat.raw)
colnames(dat.raw)
```

# Vectors for sum_score

@@

# Vectors for diff_score

>>>>

## Read in data
```{r read in data}
dat <- readRDS(file = paste0(data_path, "/four_cohorts.rds"))
dim(dat)
colnames(dat)
```

## Important vectors with collections of columns

## Vectors for sum_score

@@

## Vectors for diff_score

### 2020-09-15 PANCHANGE_descriptives.Rmd
split of scripts

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "/four_cohorts.rds")) # NEEDS CHANGING TO THE DATA SET THAT COMES OUT OF THE VARIABLE CREATION
dim(dat.raw)
colnames(dat.raw)
```

@@

```{r age_category prop total (ungrouped) plot}
#Colour palette
RAMPagepalette <- c(
  "#00D9AC","#21D9B3","#4DD9BC","#78D9C5","#8ED9C9","#A3D9CE",
  "#CFD9D7",
  "#DAD0BA","#DAC38F","#DAC38F","#DAB663" ,"#DAA937","#DA9C0C")

#Bar plot
plot.age <- 
  age_category_prop_total %>%
    mutate(age_category = fct_reorder(age_category,order)) %>%

  ggplot(aes(x = age_category)) +
  geom_bar(aes(y = n, color=age_category, fill=age_category, alpha = 1),
           stat ="summary") +
  scale_fill_manual(values =  RAMPagepalette) +
  scale_color_manual(values =  RAMPagepalette) +
  labs(x = "",
       y = "Total count",
       title = paste("Number of participants in each age category",sep=" "),
       subtitle = paste("n = ", ageN,
                        sep = ""),
       fill = "Age category") +
   guides(alpha = FALSE, color = FALSE, fill = FALSE) +
  scale_x_discrete(limits = (levels(age_category_prop_total$age_category))) +
  theme_classic2() +
   theme(axis.text.x = element_text(colour="black", size = 12,angle=45,
                                    vjust=0.5),
        axis.text.y = element_text(colour="black", size = 12))

@@

# age_category
```{r age_category per Sample prop per total}
age_category_per_Sample_prop_total <- dat %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r age_category per Sample prop cross table}
age_category_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# Ethnicity_collapsed
```{r Ethnicity_collapsed per Sample prop per total}
Ethnicity_collapsed_per_Sample_prop_total <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r Ethnicity_collapsed per Sample prop cross table}
Ethnicity_collapsed_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# Gender_collapsed
```{r Gender_collapsed per Sample prop per total}
Gender_collapsed_per_Sample_prop_total <- dat %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r Gender_collapsed per Sample prop cross table}
Gender_collapsed_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# highest_education
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# key_worker
```{r key_worker per Sample prop per total}
key_worker_per_Sample_prop_total <- dat %>%
  count(key_worker, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r key_worker per Sample prop cross table}
key_worker_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(key_worker, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

```{r gather to long format Disorder_hierarchical}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             autism_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

```{r Disorder_inc by Gender_collapsed}
Disorder_inc_by_Gender_collapsed <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Gender_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r Disorder_inc by Sample}
Disorder_inc_by_Sample <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Sample)
```

@@

```{r Disorder_inc by Ethnicity_collapsed}
Disorder_inc_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Ethnicity_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))

>>>>

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "/four_cohorts_variables.rds")) # NEEDS CHANGING TO THE DATA SET THAT COMES OUT OF THE VARIABLE CREATION
dim(dat.raw)
colnames(dat.raw)
```

## Important vectors with collections of columns

## Vectors for sum_score

```{r Vectors with sum_score_cols}
sum_score_cols <- dat.raw %>%
  select(contains("sum_score")) %>%
  select(-ends_with("raw")) %>%
  colnames()

sum_score_cols

phq.sum_score_cols <- dat.raw %>%
  select(contains("phq")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  select(-contains("8items")) %>%
  colnames()

phq.sum_score_cols

gad.sum_score_cols <- dat.raw %>%
  select(contains("gad")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

gad.sum_score_cols

ocir.sum_score_cols <- dat.raw %>%
  select(contains("ocir")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

ocir.sum_score_cols

pcl.sum_score_cols <- dat.raw %>%
  select(contains("pcl")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

pcl.sum_score_cols
```

## Vectors for diff_score

```{r Vectors with diff_score_cols}
diff_score_cols <- dat.raw %>%
  select(contains("diff_score")) %>% 
  colnames()

diff_score_cols

phq.diff_score_cols <- dat.raw %>%
  select(contains("phq")) %>%
  select(contains("diff_score")) %>% 
  colnames()

phq.diff_score_cols

gad.diff_score_cols <- dat.raw %>%
  select(contains("gad")) %>%
  select(contains("diff_score")) %>% 
  colnames()

gad.diff_score_cols

ocir.diff_score_cols <- dat.raw %>%
  select(contains("ocir")) %>%
  select(contains("diff_score")) %>% 
  colnames()

ocir.diff_score_cols
```

@@

```{r age_category prop total (ungrouped) plot}
#Colour palette
RAMPagepalette <- c(
  "#00D9AC",
  "#21D9B3",
  "#4DD9BC",
  "#78D9C5",
  "#8ED9C9",
  "#A3D9CE",
  "#CFD9D7",
  "#DAD0BA",
  "#DAC38F",
  "#DAC38F",
  "#DAB663",
  "#DAA937",
  "#DA9C0C"
  )

# Bar plot
plot.age <- 
  age_category_prop_total %>%
    mutate(age_category = fct_reorder(age_category, order)) %>%
  ggplot(aes(x = age_category)) +
  geom_bar(aes(
    y = n,
    color = age_category,
    fill = age_category,
    alpha = 1),
    stat ="identity"
    ) +
  scale_fill_manual(values =  RAMPagepalette) +
  scale_color_manual(values =  RAMPagepalette) +
  labs(x = "",
       y = "Total count",
       title = "Number of participants in each age category",
       subtitle = paste0("n = ", ageN),
       fill = "Age category") +
   guides(
     alpha = FALSE, # Display no guide for alpha
     color = FALSE, # Display no guide for color
     fill = FALSE # Display no guide for fill
     ) +
  scale_x_discrete(
    limits = (levels(age_category_prop_total$age_category))
    ) +
  theme_classic2() +
   theme(
     axis.text.x = element_text(
       colour="black",
       size = 12,
       angle = 45,
       vjust = 0.5),
     axis.text.y = element_text(
       colour="black",
       size = 12)
     )

@@

# age_category
## Proportion tables
```{r age_category per Sample prop per total}
age_category_per_Sample_prop_total <- dat %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r age_category per Sample prop cross table}
age_category_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# Ethnicity_collapsed
## Proportion tables
```{r Ethnicity_collapsed per Sample prop per total}
Ethnicity_collapsed_per_Sample_prop_total <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r Ethnicity_collapsed per Sample prop cross table}
Ethnicity_collapsed_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# Gender_collapsed
## Proportion tables
```{r Gender_collapsed per Sample prop per total}
Gender_collapsed_per_Sample_prop_total <- dat %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r Gender_collapsed per Sample prop cross table}
Gender_collapsed_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# highest_education
## Proportion tables
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

# key_worker
## Proportion tables
```{r key_worker per Sample prop per total}
key_worker_per_Sample_prop_total <- dat %>%
  count(key_worker, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r key_worker per Sample prop cross table}
key_worker_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(key_worker, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

## Create long data frame
```{r gather to long format Disorder_hierarchical}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             autism_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

## Disorders per Gender
```{r Disorder_inc by Gender_collapsed}
Disorder_inc_by_Gender_collapsed <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Gender_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Disorders per sample
```{r Disorder_inc by Sample}
Disorder_inc_by_Sample <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Sample)
```

@@

## Disorder per ethnicity
```{r Disorder_inc by Ethnicity_collapsed}
Disorder_inc_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Ethnicity_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))
